{\fontfamily{ptm}\selectfont Reading Notes}

### 2025-05-09 Notebook LaTEX/Formalized Notes/Notes 2025 03/main.tex
>w<, today's changes

        Let $F = f+ g$ where $f:\RR^n \rightarrow \overline \RR$ is differentiable and there exists $q\in \RR$ such that $g - q/2\Vert \cdot\Vert^2$ is convex.

@@

        Suppose $F = f + g$ satisfies Assumption \ref{ass:smooth-add-nonsmooth} with $\beta > 0$ and $q \in \RR$. 

@@

            \frac{q}{2}\Vert z - \bar x\Vert^2 

@@

            &\in \partial g(x^+) + \nabla f(x) + \beta(x^+ - x)

@@

            - \nabla f(x) - \beta(x^+ - x). 

@@

        Suppose $F = f + g$ satisfies Assumption \ref{ass:smooth-add-nonsmooth} such that $q = \mu_g \ge 0$, $\beta \ge L_f$. 

@@

    \section{FISTA made simple}
        We make the proofs for FISTA  with common enhancement technique available in simple proofs. 
        We showcase the theories using a generic similar triangle representations of the algorithm which tremendously simplifies the arguments. 
        \begin{assumption}[the standard FISTA setting]\label{ass:standard-fista}

@@

        Specializing Theorem \ref{thm:cnvx-pg-ineq}, we have the following lemma: 

@@

            Suppose $F = f + g$ satisfies Assumption \ref{ass:standard-fista}, define the gradient mapping for all $x \in \RR^n$
        If $f, g$ are clear in the context then we simply omit subscript and present $\mathcal G_\beta$. 

@@

        We refer to ``Generic Monotone Accelerated Proximal Gradient with line search" as ``GMAPG LS''. 
        \begin{definition}[GMAPG LS]\label{def:generic-mapg-ls}\;\\ 
            Initialize any $x_0, v_0 \in \RR^n, \alpha_0 = 1, L_0 \in \RR$ such that
                D_f\left(T_{L_0^{-1}} (y_0)\right) 
                &\le \frac{L_0}{2}\left\Vert T_{L_0^{-1}}(x_0) - y_0\right\Vert^2. 

@@

        % THEOREM | GENERIC GMAPG LS CONVERGENCE IN FXN VAL

@@

        \begin{theorem}[generic GMAPG LS convergence]\; \label{thm:gmapg-ls-convergence}\;\\

@@

            If in addition, the algorithm is initialized with $\alpha_0 = 1, x_0 = v_0 = T_{L_{0}}x_{-1} \in \text{dom}\; F$ and $x^+$ is a minimizer of $F$, then the convergence rate simplifies: 

@@

                \item $F(x_k) \le F(\tilde x_k)$ which is true by definition of GMAPG LS. 

@@

        % THEOREM | GENERIC GMAPG LS GRADIENT MAPPING CONVERGENCE

@@

        \begin{theorem}[generic GMAPG LS gradient mapping convergence]\;\label{thm:gmapg-generic-gradmap-convergence}\\

@@

            If in addition, the sequence $(\alpha_k)_{k \ge 0}$ has $\alpha_0 = 1$ and, GMAPG is initialized with $v_0=x_0 = T_{1/L_0, f, g}(x_{-1})$ for any $x_{-1} \in\RR^n$ and there exists $x^+$ which is a minimizer of $F$. 

@@

    \section{Algorithmic description of GMAPG}
        There are several components to the GMAPG algorithm. 
        This section will introduce various type of implementations that can be fitted into GMAPG in Definition \ref{def:generic-mapg-ls}. 
        \begin{algorithm}[H]
            {\small
            \begin{algorithmic}[1]\label{alg:armijo-ls}
                \STATE{\noindent
                    % \textbf{Function ArmijoLS}
                    % $(f, g, x, v, L, \alpha, \_, \_)$
                    \textbf{Function ArmijoLS:}
                    \begin{tabular}{ll}
                        $f : \RR^n \rightarrow \RR$ & Convex Lipschitz smooth\\ 
                        $g: \RR^n \rightarrow \overline \RR$  &  Convex function with Proximal operator\\ 
                        $x \in \RR^n$ & Vector\\
                        $v \in \RR^n$ & Vector\\
                        $L \in \RR$ & $L > 0$\\
                        $\alpha \in \RR$  & $\alpha \in (0, 1]$
                    \end{tabular}
                }
                \STATE{$\alpha^+ := (1/2)\left(\alpha\sqrt{\alpha^2 + 1} - \alpha^2\right)$.}
                \STATE{$y^+ := \alpha^+ v + (1 - \alpha^+)x$.}
                \STATE{$L^+ := L$. }
                \FOR{$i = 1, 2,\ldots, 53$}
                    \STATE{$L^+ := 2L^+$. }
                    \STATE{$x^+ := T_{1/L^+, f, g}(y^+)$. }
                    \IF{$D_f(x^+, y^+) \le (L^+/2)\Vert x^+ - y^+\Vert^2$}
                        \STATE{\textbf{break}}
                    \ENDIF
                    \STATE{$L^+ := 2^{i}L$}
                \ENDFOR
                \STATE{\textbf{Return:} $y^+, x^+, \alpha^+, L^+$}
                \caption{Armijo Line Search}
            \end{algorithmic}
            }
        \end{algorithm}
        \begin{algorithm}[H]\label{alg:chambolle-btls}
            {\small
            \begin{algorithmic}[1]
                \STATE{\noindent
                    % \textbf{Function ChamBT}($f, g, x, v, L, \alpha, L_{\min}, \rho$)
                    \textbf{Function ChamBT input: }
                    \begin{tabular}{ll}
                        $f:\RR^n \rightarrow \RR$ & Convex Lipschitz smooth\\ 
                        $g:\RR^n \rightarrow \overline\RR$ & Convex with Proximal Operator\\ 
                        $x\in \RR^n$ & Vector\\
                        $v\in \RR^n$ & Vector\\
                        $\alpha \in \RR$ & Vector\\
                        $L\in \RR$ & Number, $L > 0$ \\
                        $L_{\min} \in \RR$ & Number, $L_{\min} > 0$\\
                        $\rho \in \RR$ & Number, $\rho \in (0, 1)$\\
                    \end{tabular}
                }
                \STATE{$L^+ := \max(L_{\min}, \rho L)$.}
                \FOR{$i = 1, 2, \ldots, 53$}
                    \STATE{$\alpha^+ := (1/2)\left(\alpha\sqrt{\alpha^2 + L/L^+} - \alpha^2\right)$. }
                    \STATE{$y^+ := \alpha^+ v + (1 - \alpha^+) x$. }
                    \STATE{$x^+ := T_{1/L^+, f, g}(y^+)$.}
                    \IF{$2D_f(x^+, y^+) \le \Vert x^+ - y^+\Vert^2$}
                        \STATE{\textbf{break}}
                    \ENDIF
                    \STATE{$L^+ := 2^iL^+$.}
                \ENDFOR
                \STATE{\textbf{Return: }$y^+, x^+, \alpha^+, L^+$}
                \caption{Chambolle's Backtracking}
            \end{algorithmic}
            }
        \end{algorithm}
        \begin{algorithm}[H]\label{alg:beck-mono}
            \begin{algorithmic}[1]
                \STATE{\noindent
                    \textbf{Function BeckMono Input:}
                    \begin{tabular}{ll}
                        $f: \RR^n \rightarrow \RR$ & \\
                        $g: \RR^n \rightarrow \overline \RR$ & \\
                        $\tilde x \in \RR^n$ & \\
                        $x \in \RR^n$ & 
                    \end{tabular}
                \STATE{\textbf{Return: } $x^+$}

@@

        \begin{algorithm}[H]\label{alg:nes-mono}

@@

                    \textbf{Function NesMono}($f, g, x, v, L, L_{\min}, \rho, \alpha$)

@@

@@

        \begin{algorithm}[H]\label{alg:gmapg}

@@

                    \textbf{Function GMAPG} (
                        $f, g, x_{-1}, L_0, r_{\min}, \text{Exit condition: }\mathbb E_\chi$

@@

@@

            Let's consider $y_{k + 1}$ produced the GMAPG LS. 

@@

    \section{Practical enhancement from the Nesterov's Monotone Variant}
        Firstly, we will introduce some theories in for nonconvex functions. 
        This is necessary because Nesterov's monotone variants can adapt to weakly convex objective functions. 
        The following assumption will characterize the full scope of discussion in nonconvexity objective function. 
        \begin{definition}[nonconvex Nesterov's monotone scheme]\label{def:nes-monotone-scheme}

>>>>

        Let $F = f+ g$ where $f:\RR^n \rightarrow \overline \RR$ is differentiable and there exists $q\in \RR$ such that $g - \mu/2\Vert \cdot\Vert^2$ is convex.

@@

        Suppose $F = f + g$ satisfies Assumption \ref{ass:smooth-add-nonsmooth} with $\beta > 0$ and $\mu \in \RR$. 

@@

            \frac{\mu}{2}\Vert z - \bar x\Vert^2 

@@

            &\in \partial g(\bar x) + \nabla f(x) + \beta(\bar x - x)

@@

            - \nabla f(x) - \beta(\bar x - x). 

@@

        Suppose $F = f + g$ satisfies Assumption \ref{ass:smooth-add-nonsmooth} such that $\mu = \mu_g \ge 0$, $\beta \ge L_f$. 

@@

    \section{Preliminaries}
        This section introduces the full scope of the theories in our analysis. 
        \subsection{smooth plus nonsmooth weakly convex}
            \begin{assumption}[sum of weakly convex smooth and nonsmooth]\;\label{ass:sum-of-wcnvx}\\
                Let $F: \RR^n \rightarrow \overline\RR:= f + g$ such that $f, g$ satisfy 
                    \item $f$ is $L$ Lipschitz smooth and $q_f$ weakly convex. 
                    \item $g$ is $L$ Lipschitz smooth and $q_g$ weakly convex. 
            \begin{definition}[weakly convex function]\;\label{def:wcnvx-fxn}\\
                Let $f: \RR^n \rightarrow\overline \RR$ be an l.s.c proper function. 
                We define $F$ to be $q$ weakly convex if there exists $q \in \RR$ such that the function $F + q/2\Vert \cdot\Vert^2$ is a convex function and $q$ is the infimum of all such possible parameters. 
                Suppose $F = f + g$ satisfies Assumption \ref{ass:sum-of-wcnvx}, define the gradient mapping for all $x \in \RR^n$
                If $f, g$ are clear in the context then we omit subscript and present $\mathcal G_\beta$. 
            \begin{lemma}[weakly convex monotone descent]\;\label{lemma:mono-wcnvx-descent}\\
                Let $F = f + g$ satisfies Assumption \ref{ass:sum-of-wcnvx}. 
                Let $\bar x = T_{\beta, f, g}(x)$. 
                Then, for all $x \in \RR^n$, it has the following inequality: 
                $$
                \begin{aligned}
                    0 \le F(x) - F(\bar x) - (\beta - q/2 - L/2)\Vert x - \bar x\Vert^2. 
                \end{aligned}
                $$
                And descent is possible when $\beta \ge (L + q)/2$. 
                The maximum amount of descent is achieved when $\beta = L + q$. 
                Use Theorem \ref{thm:pg-ineq-wcnvx-generic}. 
        \subsection{smooth plus nonsmooth convex}
            The following assumption is strictly stronger than \ref{ass:sum-of-wcnvx}. 
            \begin{assumption}[convex smooth and nonsmooth]\label{ass:standard-fista}

@@

@@

    \section{FISTA made simple}
        We make the proofs for FISTA  with common enhancement technique available in simple proofs. 
        We showcase the theories using a generic similar triangle representations of the algorithm which tremendously simplifies the arguments.

@@

        We refer to ``Generic Monotone Accelerated Proximal Gradient with line search" as ``GMAPG''. 
        \begin{definition}[GMAPG]\label{def:generic-mapg-ls}\;\\ 
            Initialize any $x_0, v_0 \in \RR^n$. 

@@

        % THEOREM | GENERIC GMAPG CONVERGENCE IN FXN VAL

@@

        \begin{theorem}[generic GMAPG convergence]\; \label{thm:gmapg-ls-convergence}\;\\

@@

            If in addition, the algorithm is initialized with $L_0 \ge L$, $\alpha_0 = 1, x_0 = v_0 = T_{L_{0}}x_{-1} \in \text{dom}\; F$ and $x^+$ is a minimizer of $F$.
            Then, the convergence rate simplifies: 

@@

                \item $F(x_k) \le F(\tilde x_k)$ which is true by definition of GMAPG. 

@@

        % THEOREM | GENERIC GMAPG GRADIENT MAPPING CONVERGENCE

@@

        \begin{theorem}[generic GMAPG gradient mapping convergence]\;\label{thm:gmapg-generic-gradmap-convergence}\\

@@

            If in addition, the sequence $(\alpha_k)_{k \ge 0}$ has $\alpha_0 = 1$ and, GMAPG is initialized with $L_0 \ge L$, $v_0=x_0 = T_{1/L_0, f, g}(x_{-1})$ for any $x_{-1} \in\RR^n$ and there exists $x^+$ which is a minimizer of $F$. 

@@

    \section{Algorithmic description of GMAPG}
        There are several components to the GMAPG algorithm. 
        This section will introduce various type of implementations that can be fitted into GMAPG in Definition \ref{def:generic-mapg-ls}. 
            \begin{algorithmic}[1]\label{alg:armijo-ls}
                    % \textbf{Function ArmijoLS}
                    % $(f, g, x, v, L, \alpha, \_, \_)$
                    \textbf{Function ArmijoLS: }
                    \begin{tabular}{|ll}
                        $f : \RR^n \rightarrow \RR$ & Convex Lipschitz smooth\\ 
                        $g: \RR^n \rightarrow \overline \RR$  &  Convex function with Proximal operator\\ 
                        $x \in \RR^n$ & Vector\\
                        $v \in \RR^n$ & Vector\\
                        $L \in \RR$ & $L > 0$\\
                        $\alpha \in \RR$  & $\alpha \in (0, 1]$\\
                        $\cdots$ & Ignore extra inputs
                    \end{tabular}\vspace{0.5em}
                \STATE{$\alpha^+ := (1/2)\left(\alpha\sqrt{\alpha^2 + 1} - \alpha^2\right)$.}
                \STATE{$y^+ := \alpha^+ v + (1 - \alpha^+)x$.}
                \STATE{$L^+ := L$. }
                \FOR{$i = 1, 2,\ldots, 53$}
                    \STATE{$L^+ := 2L^+$. }
                    \STATE{$x^+ := T_{1/L^+, f, g}(y^+)$. }
                    \IF{$D_f(x^+, y^+) \le (L^+/2)\Vert x^+ - y^+\Vert^2$}
                        \STATE{\textbf{break}}
                    \STATE{$L^+ := 2^{i}L$}
                \STATE{\textbf{Return:} $y^+, x^+, \alpha^+, L^+$}
                \caption{Armijo Line Search}
            \begin{algorithmic}[1]\label{alg:chambolle-btls}
                    % \textbf{Function ChamBT}($f, g, x, v, L, \alpha, L_{\min}, \rho$)
                    \textbf{Function ChamBT input: }
                    \begin{tabular}{|ll}
                        $f:\RR^n \rightarrow \RR$ & Convex Lipschitz smooth\\ 
                        $g:\RR^n \rightarrow \overline\RR$ & Convex with Proximal Operator\\ 
                        $x\in \RR^n$ & Vector\\
                        $v\in \RR^n$ & Vector\\
                        $L\in \RR$ & Number, $L > 0$ \\
                        $\alpha \in \RR$ & Vector\\
                        $L_{\min} \in \RR$ & Number, $L_{\min} > 0$\\
                        $\rho \in \RR$ & Number, $\rho \in (0, 1)$\\
                    \end{tabular}\vspace{0.5em}
                \STATE{$L^+ := \max(L_{\min}, \rho L)$.}
                \FOR{$i = 1, 2, \ldots, 53$}
                    \STATE{$\alpha^+ := (1/2)\left(\alpha\sqrt{\alpha^2 + L/L^+} - \alpha^2\right)$. }
                    \STATE{$y^+ := \alpha^+ v + (1 - \alpha^+) x$. }
                    \STATE{$x^+ := T_{1/L^+, f, g}(y^+)$.}
                    \IF{$2D_f(x^+, y^+) \le \Vert x^+ - y^+\Vert^2$}
                        \STATE{\textbf{break}}
                    \STATE{$L^+ := 2^iL^+$.}
                \STATE{\textbf{Return: }$y^+, x^+, \alpha^+, L^+$}
                \caption{Chambolle's Backtracking}
        % BECK'S MONO ----------------------------------------------------------
            \begin{algorithmic}[1]\label{alg:beck-mono}
                    \textbf{Function BeckMono Inputs: }
                    \begin{tabular}{|ll}
                        $f: \RR^n \rightarrow \RR$ & \\
                        $g: \RR^n \rightarrow \overline \RR$ & \\
                        $\tilde x \in \RR^n$ & \\
                        $x \in \RR^n$ & \\
                        $\rho$ & Number $\rho \in (0, 1)$
                    \end{tabular}\vspace{0.5em}
                    $x^+ = \argmin \{(f + g)(z): z \in \{\tilde x, x\}\}$. 
                \STATE{\textbf{Return: } $x^+, \eta$}

@@

        % NES'S MONO -----------------------------------------------------------

@@

                    \textbf{Function NesMono Inputs: }
                    \begin{tabular}{|ll}
                        $f:\RR^n\rightarrow \RR$ & Lipschitz Smooth \\ 
                        $g: \RR^n \rightarrow \overline \RR$ & Weakly Convex \\ 
                        $\tilde x \in \RR^n$ & Vector \\ 
                        $x \in \RR^n$ & Vector \\
                        $\eta \in \RR$ & Number $\eta > 0$\\
                    \end{tabular}\vspace{0.5em}
                    $\hat y := \argmin\{(f + g)(z), x \in \{\tilde x, x\}\}$. 
                    $x^+ := T_{1/\eta}(\hat y)$. 
                \WHILE{$F(x^+) - F(\hat y) > -1/(2\eta)\Vert \mathcal G_{1/\eta}(\hat y)\Vert^2$}
                    \STATE{$\eta := 2\eta$. }
                \ENDWHILE
                \STATE{\textbf{return: } $x^+, \eta$ }

@@

            \end{algorithmic}\label{alg:nes-mono}

@@

        % GMAPG ----------------------------------------------------------------

@@

                    \textbf{Function GMAPG Inputs: }
                    \begin{tabular}{|ll}
                        $f:\RR^n \rightarrow \RR$ &  Weakly Convex Lipschitz Smooth \\
                        $g: \RR^n \rightarrow \overline\RR$ & Weakly Convex \\ 
                        $L\in \RR$ & $L > 0$\\ 
                        $r_{\min} $ & $r_{\min} \in (0, 1)$ \\ 
                        $\rho \in \RR$ & $\rho \in (0, 1)$ \\
                        $N_{\min}\in \N$ & $N \ge 1$ \\
                        $N \in \N$ & $N \ge N_{\min}$ \\
                        $\mathbb L_{\mathcal S}$ \\ Algorithm \ref{alg:armijo-ls} or \ref{alg:chambolle-btls}
                    \end{tabular}
                \STATE{\textbf{Return: }}

@@

            \end{algorithmic}\label{alg:gmapg}

@@

            Let's consider $y_{k + 1}$ produced the GMAPG. 

@@

    \section{Practical enhancement from the Nesterov's Monotone Variant}
        Firstly, we will introduce some theories in for nonconvex functions. 
        This is necessary because Nesterov's monotone variants can adapt to weakly convex objective functions. 
        The following assumption will characterize the full scope of discussion in nonconvexity objective function.

        \begin{definition}[nonconvex Nesterov's monotone scheme]\;\label{def:nes-monotone-scheme}\\
